---
title: "credit risk"
author: "T"
date: "1/29/2021"
output: pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(hrbrthemes)
```

This dataset was provided by the Tamkang University in Taiwan via the UCI Machine Learning Repository. The data and a description of the data can be found by following this [link](http://archive.ics.uci.edu/ml/datasets/default+of+credit+card+clients).


```{r Load Data}

credit_data <-  read_csv('credit_data.csv', col_names = TRUE)
credit_data <- credit_data %>% rename('CREDIT_AMT' = 'LIMIT_BAL', 'DEFAULT' = 'default payment next month' )

credit_data$DEFAULT <- credit_data$DEFAULT %>% factor(levels = c(0,1), labels = c('No', 'Yes'))
credit_data$SEX <- credit_data$SEX %>% factor(levels = c(1,2), labels = c('Male', 'Female'))
credit_data$EDUCATION <- as.factor(credit_data$EDUCATION)
credit_data$MARRIAGE <- as.factor(credit_data$MARRIAGE)

glimpse(credit_data)


```


```{r Credit line Distribution}
dens_scale = 5e-6 / 7000

credit_data %>% ggplot(aes(x = CREDIT_AMT)) +
  geom_histogram(color = 'black', fill = rgb(0.2, 0.6, 0.9, 1), alpha = 0.3) +
  geom_density(aes(y = ..scaled.. * 7000), color = 'red') +
  scale_y_continuous(
    name = 'Counts', sec.axis = sec_axis(~.*dens_scale, name = 'Density')
  ) + 
    theme_ipsum() +
  ggtitle('') + 
  xlab('Amount of Credit') +
  theme(
    axis.title.x = element_text(size = 11, hjust = 0.5),
    axis.title.y = element_text(color = rgb(0.2, 0.6, 0.9, 1), size = 13), 
    axis.title.y.right = element_text(color = 'red', size = 13)
  ) 
  
```


## Who defaulted tho?
```{r Defaulted}
# Number of Loans Defaulted
credit_data %>% ggplot(aes(x = DEFAULT)) + 
  geom_bar(aes(fill = DEFAULT), show.legend = FALSE) +
  xlab('Default on Loan') + 
  ylab('Number of Applicants') + 
  ggtitle('Loan Defaul Rate') + 
  geom_text(stat = 'count', aes(label = ..count..), position = position_stack(vjust= 0.5)) 

# By Age
credit_data %>% ggplot(aes(x = DEFAULT, y = AGE)) + geom_boxplot(aes(fill = DEFAULT), show.legend = FALSE)
t.test(credit_data$AGE ~ credit_data$DEFAULT)

# By Gender
credit_data %>% ggplot(aes(x = DEFAULT)) + geom_bar(aes(fill = SEX), position = 'dodge')
chisq.test(credit_data$DEFAULT, credit_data$SEX)


# Defaulted by Gender
credit_data %>% 
  filter(DEFAULT == 'Yes') %>% 
  ggplot(aes(x = DEFAULT)) + 
    geom_bar(aes(fill = SEX), position = 'dodge')+
    xlab('Defaulted') + 
    scale_x_discrete(labels = '') +
    ylab('Number of Customers') +
    ggtitle('Number of Customers who Defaulted on Loan') 

def_mf <- credit_data %>% filter(DEFAULT == "Yes") %>% select(SEX) %>% count(SEX)
prop.test(male_def <- def_mf[1,2][[1]], colSums(def_mf[2]))
    # Looks lke women tend to default on loans more often. Digging deepeer reveals this isnt true. 



# By Credit Amount
credit_data %>% 
  group_by(DEFAULT) %>% 
  summarise(avg_cred_amt = mean(CREDIT_AMT)) %>% 
  ggplot(aes(DEFAULT, avg_cred_amt)) +
    geom_bar(stat = 'identity', aes(fill = DEFAULT), show.legend = FALSE) +
    xlab('Defaulted on Loan') + 
    ylab('Average Amount of Loan') +
    ggtitle('Average Loan by Default Status') 

credit_data %>% 
  ggplot(aes(DEFAULT, CREDIT_AMT)) +
    geom_violin(aes(fill = DEFAULT), show.legend = FALSE) + 
    xlab('Defaulted on Loan') + 
    ylab('Amount of Loan') +
    ggtitle('Loan Distribution by Default Status') 
    # Defaulted loans were worth less, why? (potentially riskier clients?)



# By Education
credit_data %>% ggplot(aes(EDUCATION)) + 
  geom_bar(aes(fill = EDUCATION), show.legend = FALSE)

credit_data %>% 
  ggplot(aes(EDUCATION, fill = DEFAULT)) +
    geom_bar(position = 'dodge')

ed_def <- credit_data %>% 
  group_by(EDUCATION, DEFAULT) %>% 
  summarise(count = n()) %>% 
  mutate(perc = count/sum(count))
ed_def %>% 
  ggplot(aes(EDUCATION, perc, fill = DEFAULT)) +
    geom_bar(stat = "identity", position = 'dodge')



# By Marriage
credit_data %>% ggplot(aes(MARRIAGE)) +
  geom_bar(aes(fill = MARRIAGE), show.legend = FALSE)

credit_data %>% ggplot(aes(MARRIAGE)) +
  geom_bar(aes(fill = DEFAULT), position = 'dodge')

mar_def <- credit_data %>% 
  group_by(MARRIAGE, DEFAULT) %>% 
  summarize(count = n()) %>% 
  mutate(perc = count/sum(count))
mar_def %>% ggplot(aes(MARRIAGE)) +
  geom_bar(aes(y = perc, fill = DEFAULT), stat = 'identity', position = 'dodge')

```

Gives the impression that more females defaulted on their loans, but in reality there are more females overall that received loans. 



## Break down by Gender
Inspecting the disparity in defaults between the women and men that were approved for a loan but defaulted.

```{r Gender}
credit_data %>% ggplot(aes(SEX)) + geom_bar(aes(fill = SEX))

def_by_gender <- credit_data %>% group_by(SEX, DEFAULT) %>% summarize(count = n()) %>% mutate(perc = count/sum(count))

brks = c(0, 0.25, 0.50, 0.75, 1)
def_by_gender %>% ggplot(aes(x = DEFAULT, y = perc)) +
  geom_bar(aes(fill = SEX), position = 'dodge', stat = "identity") +
  xlab('Defaulted on Loan') + 
  ylab('Percent') +
  ggtitle('Percent of Customers who Defaulted \n on Loan by Gender') +
  scale_y_continuous(breaks = brks, labels = scales::percent(brks))

def_by_gender %>% filter(DEFAULT == 'Yes') %>% select(perc)


credit_data %>% 
  filter(DEFAULT == 'Yes') %>% 
  ggplot(aes(x = DEFAULT)) + 
    geom_bar(aes(fill = SEX), position = 'fill') +
    xlab('Defaulted') + 
    scale_x_discrete(labels = '') +
    ylab('Number of Customers') +
    ggtitle('Number of Customers who Defaulted on Loan')

```


  
